Schema Evolution in Interactive Programming Systems

### 2024-08-28 paper.tex
clean up first mention of local-first

The Extract Entity challenge in the non-local case (for local-first software) is similar to the
local live programming challenge except that the context is an application that offers similar
functionality with peer-to-peer collaboration. In this context the schema change does not need
to be performed interactively -- a developer can take some time to build, test, and package a
solution, perhaps using an API or DSL. The hard part comes when the schema change is to
be deployed across all the replicas. This deployment must synchronize code and data upgrades
(or decouple them as in Cambria~\cite{Cambria}). The deployment must also deal with
migrating ``in flight'' operations so that all replicas converge on the same state
without data loss, and ideally without centralized coordination. A radical approach
could try to incorporate schema change operations into the underlying datastore
itself (possibly a CRDT) and integrate code as well, but layered solutions are welcome too.

>>>>

We can also consider the Extract Entity challenge in a local-first ~\cite{localfirst} case, where the context is an application that offers peer-to-peer collaboration. In this context the schema change does not need
to be performed interactively -- a developer can take some time to build, test, and package a
solution, perhaps using an API or DSL. The hard part comes when the schema change is to
be deployed across all the replicas. This deployment must synchronize code and data upgrades
(or decouple them as in Cambria~\cite{Cambria}). The deployment must also deal with
migrating ``in flight'' operations so that all replicas converge on the same state
without data loss, and ideally without centralized coordination. A radical approach
could try to incorporate schema change operations into the underlying datastore
itself (possibly a CRDT) and integrate code as well, but layered solutions are welcome too.
